Compare with SH HA paper

### 2017-01-18 Code/Run_model_on_SH_data/Prep SH HA data for our analysis.Rmd

```{r}
library(foreign)
library(dplyr)

dtotal <- read.table("/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total-nhisp.txt", header = F, sep="\t")

names(dtotal)
#no names affixed to dataset. But the names are stored in the data dictionary file found at:
#/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total-nhisp.dic

@@

```{r recode_variables}
dtotal <- dtotal %>%
  mutate(State2 = factor(State, levels = c(0:50), labels = c("Alabama", "Alaska", "Arizona", "Arkansas","California",
                                                             "Colorado", "Connecticut", "Delaware", "Washington DC",
                                                             "Florida", "Georgia", "Hawaii","Idaho", "Illinois", 
                                                             "Indiana","Iowa", "Kansas", "Kentucky", "Louisiana", 
                                                             "Maine", "Maryland", "Massachusetts", "Michigan", 
                                                             "Minnesota", "Mississippi","Missouri", "Montana", 
                                                             "Nebraska", "Nevada", "New Hampshire", "New Jersey",
                                                             "New Mexico", "New York","North Carolina","North Dakota",
                                                             "Ohio", "Oklahoma","Oregon", "Pennsylvania", 
                                                             "Rhode Island", "South Carolina", "South Dakota", 
                                                             "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
                                                             "Washington", "West Virginia", "Wisconsin", "Wyoming")), 
         Age2 = factor(Age, levels = c(0:19), labels = c("<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                                         "15-19 years", "20-24 years", "25-29 years", "30-34 years",
                                                         "35-39 years", "40-44 years", "45-49 years", "50-54 years", 
                                                         "55-59 years", "60-64 years", "65-69 years", "70-74 years", 
                                                         "75-79 years", "80-84 years","85+ years", "Unknown"), ordered = T),
         Age3 = 0*(Age2 == "<1 year") + 1*(Age2 == "1-4 years") + 5*(Age2 == "5-9 years") +
                10*(Age2 == "10-14 years") + 15*(Age2 == "15-19 years") + 20*(Age2 == "20-24 years") + 
                25*(Age2 == "25-29 years") + 30*(Age2 == "30-34 years") + 35*(Age2 == "35-39 years") +
                40*(Age2 == "40-44 years") + 45*(Age2 == "45-49 years") + 50*(Age2 == "50-54 years") + 
                55*(Age2 == "55-59 years") + 60*(Age2 == "60-64 years") + 65*(Age2 == "65-69 years") +   
                70*(Age2 == "70-74 years") + 75*(Age2 == "75-79 years") + 80*(Age2 == "80-84 years") + 
                85*(Age2 == "85+ years"),
         nx = 1*(Age2 == "<1 year") + 4*(Age2 == "1-4 years") + 5*(Age2 != "<1 year" & Age2 != "1-4 years"),
         Sex2 = factor(Sex, levels = c(0, 1), labels = c("Male", "Female")),
         Race2 = factor(Race, levels = c(0, 1), labels = c("White", "Black")),
         Year2 = factor(Year, levels = c(0:19), labels = as.character(1990:2009)),
         Year3 = as.numeric(as.character(Year2)),
         RaceSex = interaction(Race2, Sex2, sep = " "),
         StateYearSex = interaction(State2, Year2, Sex2),
         StateYearRaceSex = interaction(State2, Year2, Race2, Sex2))
```

@@

#Count of death is always equal to 0 for unknown age groups
table(dtotal$Age, dtotal$Count==0)

@@

@@

>>>>

```{r}
library(foreign)
library(dplyr)
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")

dtotal <- read.table("/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total.txt", header = F, sep="\t", stringsAsFactors = F, colClasses = c(rep("numeric", 8)),
                     na.strings = ".")

names(dtotal)
#no names affixed to dataset. But the names are stored in the data dictionary file found at:
#/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total.dic

@@

```{r recode_variables}
dtotal <- dtotal %>%
  mutate(State2 = factor(State, levels = c(0:50), labels = c("Alabama", "Alaska", "Arizona", "Arkansas","California",
                                                             "Colorado", "Connecticut", "Delaware", "Washington DC",
                                                             "Florida", "Georgia", "Hawaii","Idaho", "Illinois", 
                                                             "Indiana","Iowa", "Kansas", "Kentucky", "Louisiana", 
                                                             "Maine", "Maryland", "Massachusetts", "Michigan", 
                                                             "Minnesota", "Mississippi","Missouri", "Montana", 
                                                             "Nebraska", "Nevada", "New Hampshire", "New Jersey",
                                                             "New Mexico", "New York","North Carolina","North Dakota",
                                                             "Ohio", "Oklahoma","Oregon", "Pennsylvania", 
                                                             "Rhode Island", "South Carolina", "South Dakota", 
                                                             "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
                                                             "Washington", "West Virginia", "Wisconsin", "Wyoming")), 
         Age2 = factor(Age, levels = c(0:19), labels = c("<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                                         "15-19 years", "20-24 years", "25-29 years", "30-34 years",
                                                         "35-39 years", "40-44 years", "45-49 years", "50-54 years", 
                                                         "55-59 years", "60-64 years", "65-69 years", "70-74 years", 
                                                         "75-79 years", "80-84 years","85+ years", "Unknown"), ordered = T),
         Age3 = 0*(Age2 == "<1 year") + 1*(Age2 == "1-4 years") + 5*(Age2 == "5-9 years") +
                10*(Age2 == "10-14 years") + 15*(Age2 == "15-19 years") + 20*(Age2 == "20-24 years") + 
                25*(Age2 == "25-29 years") + 30*(Age2 == "30-34 years") + 35*(Age2 == "35-39 years") +
                40*(Age2 == "40-44 years") + 45*(Age2 == "45-49 years") + 50*(Age2 == "50-54 years") + 
                55*(Age2 == "55-59 years") + 60*(Age2 == "60-64 years") + 65*(Age2 == "65-69 years") +   
                70*(Age2 == "70-74 years") + 75*(Age2 == "75-79 years") + 80*(Age2 == "80-84 years") + 
                85*(Age2 == "85+ years"),
         nx = 1*(Age2 == "<1 year") + 4*(Age2 == "1-4 years") + 5*(Age2 != "<1 year" & Age2 != "1-4 years"),
         Sex2 = factor(Sex, levels = c(0, 1), labels = c("Male", "Female")),
         Race2 = factor(Race, levels = c(0, 1), labels = c("White", "Black")),
         Year2 = factor(Year, levels = c(0:19), labels = as.character(1990:2009)),
         Year3 = as.numeric(as.character(Year2)),
         RaceSex = interaction(Race2, Sex2, sep = " "),
         StateYearSex = interaction(State2, Year2, Sex2),
         StateYearRaceSex = interaction(State2, Year2, Race2, Sex2),
         Crude_Rate = Count/Population)
```

@@

#no rows with the unknown age group (guess: was removed when pulling the data from CDC wonder)
table(dtotal$Age, dtotal$Count==0)

@@

excluded.states <- c("Idaho", "Montana", "North Dakota", "South Dakota", "Vermont")
included.states <- setdiff(levels(dtotal$State2), excluded.states)
dtotal <- dtotal %>% filter(State2 %in% included.states)

@@

```{r}
prop.table(table(is.na(dtotal$Count)))

prop.table(table(is.na(dtotal$Count[dtotal$State2 == "New York"])))
prop.table(table(is.na(dtotal$Count[dtotal$State2 == "Georgia"])))
prop.table(table(is.na(dtotal$Count[dtotal$State2 == "Florida"])))

test.le.fx <- function(data, state, sex, race, year) {
  test.data <- data[data["State2"] == state & data["Sex2"] == sex & data["Race2"] == race & data["Year3"] == year, ]
  lt <- life.table(data = test.data, num.ages.in.group = "nx",
             death.counts = "Count", population.counts = "Population")
  return(round(lt$e_x[1], 1))
}

test.le.fx(dtotal, "Florida", "Male", "White", 1990)
test.le.fx(dtotal, "Florida", "Male", "White", 2009)

test.le.fx(dtotal, "Georgia", "Male", "White", 1990)
test.le.fx(dtotal, "Georgia", "Male", "White", 2009)

test.le.fx(dtotal, "New York", "Male", "White", 1990)
test.le.fx(dtotal, "New York", "Male", "White", 2009)
```
